---
output: github_document
---

<!-- badges: start -->
  [![R build status](https://github.com/FredHutch/VISCfunctions/workflows/R-CMD-check/badge.svg)](https://github.com/FredHutch/VISCfunctions/actions)
[![Codecov test coverage](https://codecov.io/gh/FredHutch/VISCfunctions/graph/badge.svg)](https://app.codecov.io/gh/FredHutch/VISCfunctions)
  [![License:MIT](https://img.shields.io/badge/License-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
<!-- badges: end -->

  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# VISCfunctions

The package provides functions for common tasks for creating statistical reports at Vaccine Immunology Statistical Center (VISC).  The goal of VISCfunctions is to:

* provide standard statistical testing and estimating functions
* help format output for PDF and Word reports
* provide frequently used utility functions

for VISC statisticians and programmers at Fred Hutch.

# Installation

The package is available on the Fred Hutch organization GitHub page.

```{r eval=FALSE}
remotes::install_github("FredHutch/VISCfunctions")

# Use the build_vignettes parameter to access the vignette:
remotes::install_github("FredHutch/VISCfunctions", build_vignettes = TRUE)
vignette("Overview")
```

# Overview

Below is an overview of the currently available functions in VISCfunctions.

## Statistical testing and estimates

Compare two groups, return a p-value:

* `two_sample_bin_test()` for a Barnard, Fisher's Exact, Chi-Square or McNemar test.
* `two_samp_cont_test()` for a t.test (paired or unpaired), Wilcox Rank-Sum, or Wilcox Signed-Rank test.
* `cor_test()` for a Spearman, Pearson, or Kendall correlation test.

Make all pairwise comparisons of a grouping variable (or any categorical variable), return descriptive statistics and p-values:

* `pairwise_test_bin()` for a Barnard, Fisher's Exact, Chi-Square or McNemar test.
* `pairwise_test_cont()` for a t.test (paired or unpaired), Wilcox Rank-Sum, or Wilcox Signed-Rank test.
* `cor_test_pairs()` for a Spearman, Pearson, or Kendall correlation test.

Estimate binomial confidence intervals for a binary vector:

* `binom_ci()`

## Formatting output

* `paste_tbl_grp()` to paste together information (usually descriptive statistics) from two groups.
* `pretty_pvalues()` to round and format p-values.
* `stat_paste()` to combine and format values, such as:
    + Mean (sd)
    + Median [min, max]
    + Estimate (SE of Estimate)
    + Estimate (95% CI Lower Bound, Upper Bound)
    + Estimate/Statistic (p value)
* `escape()` to protect control characters in a string for use in a latex table or caption
* `collapse_group_row()` as an alternative to `kableExtra::collapse_rows()` for long tables in latex

## Survival and Magnitude Breadth

* `create_step_curve()` to create survival probabilities and generate a risk table
* `mb_results()` to create step curve info for magnitude breadth (MB) plots and AUC-MB

## Utility functions

* `round_away_0()` is an alternative to the `round()` function to properly perform mathematical rounding.
* `escape()` is used to inserts a "\\" in front of values, which is needed for Latex.
* `get_full_name()` looks up a username from Fred Hutch ID.
* `get_session_info()` creates a a data frame with session information.


# Vignette

For more information, browse the vignette (`browseVignettes("VISCfunctions")`).

# Contribute to this package

See our [contibuting guide](CONTRIBUTING.md) to learn more about how you can contribute to this package.
