---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# VISCfunctions

The package provides functions for common tasks for creating statistical reports at Vaccine Immunology Statistical Center (VISC).  The goal of VISCfunctions is to:

* provide standard statistical testing and estimating functions
* help format output for PDF and Word reports
* provide frequently used utility functions

for VISC statisticians and programmers at Fred Hutch.

# Installation

The package is available on the Fred Hutch organization GitHub page.

```{r eval=FALSE}
remotes::install_github("FredHutch/VISCfunctions")

# Use the build_vignettes parameter to access the vignette:
remotes::install_github("FredHutch/VISCfunctions", build_vignettes = TRUE)
vignette("Overview")
```

# Overview

Below is an overview of the currently available functions in VISCfunctions.

## Statistical testing and estimates

Compare two groups, return a p-value:

* `two_sample_bin_test()` for a Barnard, Fisher's Exact, Chi-Square or McNemar test.
* `two_samp_cont_test()` for a t.test (paired or unpaired), Wilcox Rank-Sum, or Wilcox Signed-Rank test.

Make all pairwise comparisons of a grouping variable (or any categorical variable), return descriptive statistics and p-values:

* `pairwise_test_bin()` for a Barnard, Fisher's Exact, Chi-Square or McNemar test.
* `pairwise_test_cont()` for a t.test (paired or unpaired), Wilcox Rank-Sum, or Wilcox Signed-Rank test.

Estimate wilson confidence intervals for a binary vector:

* `wilson_ci()`

## Formatting output

* `paste_tbl_grp()` to paste together information (usually descriptive statistics) from two groups.
* `pretty_pvalues()` to round and format p-values.
* `stat_paste()` to combine and format values, such as:
    + Mean (sd)
    + Median [min, max]
    + Estimate (SE of Estimate)
    + Estimate (95% CI Lower Bound, Upper Bound)
    + Estimate/Statistic (p value)

## Utility functions

* `round_away_0()` is an alternative to the `round()` function to properly perform mathematical rounding.
* `escape()` is used to inserts a "\\" in front of values, which is needed for Latex.
* `get_full_name()` looks up a username from Fred Hutch ID.
* `get_session_info()` creates a a data frame with session information.


# Vignette

For more information, browse the vignette (`browseVignettes("VISCfunctions")`).

# Contribute to this package

See our [contibuting guide](CONTRIBUTING.md) to learn more about how you can contribute to this package.
