---
title: "Introduction to VISCfunctions"
author: "Monica Gerber"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
vignette: >
  %\VignetteIndexEntry{Introduction to VISCfunctions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<!-- I created this vignette with usethis::use_vignette. -->
<!-- There's also good guidance here: http://r-pkgs.had.co.nz/vignettes.html -->


VISCfunctions aims to provide a function for common statistical tests and data manipulation tasks performed at VISC.


## Data: exampleData_BAMA

```{r}
library(VISCfunctions)
data(exampleData_BAMA)
str(exampleData_BAMA)
```

## Satistical tests

### P-values from a two-sample test for continuous data with `two_samp_cont_test`

The `two_samp_cont_test` function allows you to conduct tests of independence between two continuous variables, and then returns the p-value for that test. This function uses the `coin` package to implement these tests.

For example, we can test the null hypothesis, *magnitude~group1~* is equal to *magnitude~group2~* at visit 1 for the gp41 antigen.

```{r}
visitno1_gp41 <- subset(exampleData_BAMA, visitno == 1 & antigen == "gp41")

# using a t.test
two_samp_cont_test(
  x = visitno1_gp41$magnitude, 
  y = visitno1_gp41$group,
  method = "t.test",
  paired = FALSE,
  )

# using a Wilcoxon rank-sum test
two_samp_cont_test(
  x = visitno1_gp41$magnitude, 
  y = visitno1_gp41$group,
  method = "wilcox",
  paired = FALSE,
  )

```

Note that if this was a paired sample, you can use `paired = TRUE` for a paired t-test or a Wilcoxon signed-rank test. Use `verbose = TRUE` to see errors.

### P-values from a two-sample test for binary data with `two_samp_bin_test`

The `two_samp_bin_test` function allows you to conduct tests of independence between two binary variables, and then returns the p-value for that test. This function uses base R to implement Fisher's, chi-square, and McNemar's tests. It uses the `Exact` package to implement the Barnard's test.

For example, we can test the null hypothesis, the response rate in group 1 is equal to the response rate in group 2 at visit 1 for the gp41 antigen.

```{r}
# using barnard's test
# default is two.sided test
two_samp_bin_test(
  x = visitno1_gp41$response,
  y = visitno1_gp41$group,
  method = "barnard"
)
```


### Test multiple groups with `pairwise_test_cont`

If there are many groups in a study, it is often necessary to do pairwise comparisons. For continuous data, use the `pairwise_test_cont` function. This function outputs the median (min, max) as well as the p-value for each comparison. To make group comparisons for each antigen and visit number, do:

```{r}
library(dplyr)
library(tidyr)
library(purrr)

exampleData_BAMA %>%
  group_by(antigen, visitno) %>%
  nest() %>%
  mutate(
    pairwise_test = map(
      data,
      ~pairwise_test_cont(
        x = .$magnitude,
        group = .$group,
        paired = FALSE, 
        method = 'wilcox', 
        alternative = "less", 
        digits = 3, 
        num_needed_for_test = 3
        )
      )
  ) %>%
  select(-data) %>%
  unnest()
```

Note that the `num_needed_for_test` argument is useful for statistical tests that require at least 3 responders. The "Comparison" column will reflect the type of `alternative` hypothesis: two-sided, less than, or greater than. The `digits` and `trailing_zeros` arguements specify rounding and formatting of values. 

## Rounding and formatting data

### Round and format values with `round_away_0`

For tables in PT reports, rounding is often necessary.

```{r}
exampleData_BAMA %>%
  group_by(antigen, visitno, group) %>%
  summarise(
    mean = mean(magnitude, na.rm = TRUE) %>% 
      round_away_0(., digits = 1, trailing_zeros = TRUE)
  )
```

Note that `round_away_0` will return a character vector if `trailing_zeros = TRUE`.

### Round and format p-values with `pretty_pvalues`.

P-values need to be formatted for PT reports. The `pretty_pvalues` function will round a vector of p-values, and use the "<" symbol for p-values below the specified digit. It returns a character vector.

```{r}
exampleData_BAMA %>%
  group_by(antigen, visitno) %>%
  nest() %>%
  mutate(
    pval_group_diff = map_chr(
      data,
      ~two_samp_cont_test(
        .$magnitude,
        .$group,
        method = "wilcox"
      ) %>% 
        pretty_pvalues(., digits = 2)
    )
  ) %>%
  select(-data)
```


### Format results for tables with `stat_paste`

To create presentation-worthy tables, it's often necessary to combine multiple variables in the same column. For example, the median (min, max) for data is often presented:

```{r}
exampleData_BAMA %>%
  group_by(antigen, visitno, group) %>%
  summarise(
    median = median(magnitude, na.rm = TRUE),
    min = min(magnitude, na.rm = TRUE),
    max = max(magnitude, na.rm = TRUE)
  ) %>%
  mutate(
    med_min_max = stat_paste(median, min, max, digits = 1)
  ) %>%
  select(antigen:group, med_min_max)
```

Note that rounding can also be done with this function. You can also specify the "{}", "[]", or "|" symbols instead of "()" with the argument `bound_char = "{"`, etc.

### Format results to compare two groups with `paste_tbl_grp`

To easily visually compare results from two groups, paste together variables from both groups into a single column with the `paste_tbl_grp` function. By default, `paste_tbl_grp` will paste together data from two groups for median_min_max or mean_sd if this exists in the dataset. 

```{r}
exampleData_BAMA %>%
  mutate(group = paste0("Group", group)) %>%
  group_by(group, visitno, antigen) %>%
  summarise_at("magnitude", funs(mean, sd, median, min, max)) %>%
  gather(variable, value, -(group:antigen)) %>% # these three chains create a wide dataset
  unite(temp, group, variable) %>%
  spread(temp, value) %>%
  mutate(Group1 = "Group 1", Group2 = "Group 2") %>%
  paste_tbl_grp() %>%
  select(visitno:antigen, median_min_max_comparison) %>%
  head()
```

Specify variables to paste with `vars_to_paste` and group names with `first_name` and `second_name`. If you want to replace "vs." with a different character string, use `sep_val`.


<!-- note: I didn't include `update_visc_data` - is for updating data packages? I think this needs to be updated now that we're using github.com. -->




