---
title: "Introduction to VISCfunctions"
author: "Monica Gerber"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to VISCfunctions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


VISCfunctions aims to provide a function for common statistical tests and data manipulation tasks performed at VISC:


## Data: exampleData_BAMA

```{r}
data(exampleData_BAMA)
str(exampleData_BAMA)
```

## Get p-values from statistical tests

### Two-sample test for continuous data with `two_samp_cont_test`

The `two_samp_cont_test` function allows you to conduct tests of independence between two continuous variables using, and then returns the p-value for that test. This function uses the `coin` package to implement these tests.

For example, we can test the null hypothesis, *magnitude~group1~* is equal to *magnitude~group2~* at visit 1 for the gp41 antigen.

```{r}
visitno1_gp41 <- subset(exampleData_BAMA, visitno == 1 & antigen == "gp41")

# using a t.test
two_samp_cont_test(
  x = visitno1_gp41$magnitude, 
  y = visitno1_gp41$group,
  method = "t.test",
  paired = FALSE,
  )

# using a Wilcoxon rank-sum test
two_samp_cont_test(
  x = visitno1_gp41$magnitude, 
  y = visitno1_gp41$group,
  method = "wilcox",
  paired = FALSE,
  )

```

Note that if this was a paired sample, you can use `paired = TRUE` for a paired t-test or a Wilcoxon signed-rank test. Use `verbose = TRUE` to see errors.

### Two-sample test for binary data with `two_samp_bin_test`

The `two_samp_bin_test` allows you to conduct tests of independence between two binary variables, and then returns the p-value for that test. This function uses base R to implement Fisher's, chi-square, and McNemar's tests. It uses the `Exact` package to implement the Barnard's test.

For example, we can test the null hypothesis, the response rate in group 1 is equal to the response rate in group 2 at visit 1 for the gp41 antigen.

```{r}
# using barnard's test
# default is two.sided test
two_samp_bin_test(
  x = visitno1_gp41$response,
  y = visitno1_gp41$group,
  method = "barnard"
)
```


### Test multiple groups with `pairwise_test_cont`

If there are many groups in a study, it is often necessary to do pairwise comparisons. For continuous data, use the `pairwise_test_cont` function. This function outputs the sample statistics as well as the p-value for each comparison. To make group comparisons for each antigen and visit number, do:

```{r}
library(dplyr)
library(tidyr)
library(purrr)

exampleData_BAMA %>%
  group_by(antigen, visitno) %>%
  nest() %>%
  mutate(
    pairwise_test = map(
      data,
      ~pairwise_test_cont(
        x = .$magnitude,
        group = .$group,
        paired = FALSE, 
        method = 'wilcox', 
        alternative = "less", 
        digits = 3, 
        num_needed_for_test = 3
        )
      )
  ) %>%
  select(-data) %>%
  unnest()
```


## Rounding and formatting data

### `round_away_0`

### `pretty_pvalues`

### `stat_paste`

### `paste_tble_grp`


## Other


### `update_visc_data`






Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))

