---
title: "Overview and Introduction to VISCfunctions Package"
author: "Jimmy Fulp, Monica Gerber, and Heather Bouzek"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    df_print: kable
    highlight: tango
    toc: true
    toc_depth: 2
    number_sections: true
    keep_tex: true
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
vignette: >
  %\VignetteIndexEntry{Overview and Introduction to VISCfunctions Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

\hypersetup{colorlinks=true, linkcolor=blue}

\listoftables

\clearpage


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

options(
  xtable.comment = FALSE,
  datatable.verbose = FALSE,
  scipen = 10,
  knitr.kable.NA = '',
  knitr.table.format = 'latex',
  kableExtra.latex.load_packages = FALSE
)

```



# Overview

VISCfunctions is a collection of useful functions designed to assist in the
analysis and creation of professional reports. The current VISCfunctions package
can be broken down to the following sections:

- Testing and Estimate Functions
  - two_samp_bin_test
  - two_samp_cont_test
  - cor_test
  - binom_ci
  - trapz_sorted
- Fancy Output Functions
  - pretty_pvalues
  - stat_paste
  - paste_tbl_grp
  - escape
  - collapse_group_row
- Pairwise Comparison Functions
  - pairwise_test_bin
  - pairwise_test_cont
  - cor_test_pairs
- Survival and Magnitude Breadth
  - create_step_curve
  - mb_results
- Utility Functions
  - round_away_0
  - get_session_info
  - get_full_name
- Example Datasets
  - exampleData_BAMA
  - exampleData_NAb
  - exampleData_ICS
  - CAVD812_mAB


## Getting Started


Code to initially install the VISCfunctions package:

```{r, echo=TRUE, eval=FALSE}

# Installing VISCfunctions from GitHub
remotes::install_github("FredHutch/VISCfunctions")
```

Code to load in VISCfunctions and start using:

```{r, echo=TRUE, results='hide', message=FALSE}

# Loading VISCfunctions
library(VISCfunctions)

# Loading dplyr for pipe
library(dplyr)
```

\clearpage

## VISCtemplates Package

The [VISCtemplates](https://github.com/FredHutch/VISCtemplates) package makes
extensive use of the `VISCfunctions` package, and is a great way get started
making professional statistical reports.

Code to initially download VISCtemplates package:

```{r, echo=TRUE, eval=FALSE}

# Installing VISCfunctions from GitHub
remotes::install_github("FredHutch/VISCtemplates")
```

Once installed, in RStudio go to **File -> New File -> R Markdown -> From
Template -> VISC PT Report (Generic)** to start a new Markdown report using the
template. Within the template there is code to load and make use of most of the
`VISCfunctions` functionality.

## Example Datasets

The `exampleData_BAMA`, `exampleData_NAb`, `exampleData_ICS`, and `CAVD812_mAB`
datasets are assay example datasets used throughout this vignette and most
examples in the `VISCfunctions` documentation. All example datasets have
associated documentation that can be viewed using `?` (i.e `?exampleData_BAMA`).

```{r, echo=TRUE}

# Loading in example datasets
data("exampleData_BAMA", "exampleData_NAb", "exampleData_ICS", "CAVD812_mAB")

# Quick view of dataset
str(exampleData_BAMA)
```


```{r, echo=FALSE}
# Loading in example datasets
exampleData_BAMA <- 
  exampleData_BAMA %>% 
  mutate(antigen = gsub('_', ' ', antigen))
```


# Testing and Estimate Functions{#testing-and-estimate-functions}

There are currently three testing functions performing the appropriate
statistical tests, depending on the data and options, returning a p value.

There is also an estimate function for getting binary confidence intervals
(`binom_ci()`).

`trapz_sorted()` is used to create an estimate for the area under a curve while
making sure the x-axis is sorted so that both the x-axis and the area are
increasing and positive, respectively.

```{r, echo=TRUE}

# Making Testing Dataset
testing_data <- exampleData_BAMA %>%
  dplyr::filter(antigen == 'A1.con.env03 140 CF' & visitno == 1)

```


\clearpage

## Comparing Two Groups (Binary Variable) for a Binary Variable 

`two_samp_bin_test()` is used for comparing a binary variable to a binary (two
group) variable, with options for Barnard, Fisher's Exact, Chi-Sq, and McNemar
tests. For the Barnard test specifically, there are many model options that can
be set that get passed to the `Exact::exact.test()` function.


```{r, echo=TRUE, warning=FALSE}

table(testing_data$response, testing_data$group)

# Barnard Method
two_samp_bin_test(
  x = testing_data$response,
  y = testing_data$group,
  method = 'barnard',
  alternative = 'two.sided'
)

# Santner and Snell Variation
two_samp_bin_test(
  x = testing_data$response,
  y = testing_data$group,
  method = 'barnard',
  barnard_method = 'santner and snell',
  alternative = 'two.sided')

# Calling test multiple times
exampleData_BAMA %>% 
  group_by(antigen, visitno) %>% 
  filter(visitno != 0) %>% 
  summarise(p_val = two_samp_bin_test(response, group, method = 'barnard'))
```


\clearpage

## Comparing Two Groups (Binary Variable) for a Continuous Variable 

`two_samp_cont_test()` is used for comparing a continuous variable to a binary
(two group) variable, with parametric (t.test) and non-parametric (Wilcox
Rank-Sum) options. Also paired data is allowed, where there are parametric
(paired t.test) and non-parametric (Wilcox Signed-Rank) options.



```{r, echo=TRUE, warning=FALSE}

by(testing_data$magnitude, testing_data$group, summary)

two_samp_cont_test(
  x = testing_data$magnitude,
  y = testing_data$group,
  method = 'wilcox',
  alternative = 'two.sided'
)

# Calling test multiple times
exampleData_BAMA %>% 
  group_by(antigen, visitno) %>% 
  filter(visitno != 0) %>% 
  summarise(p_val = two_samp_cont_test(magnitude, group, method = 'wilcox'))
```


\clearpage

## Comparing Two Continuous Variables (Correlation)

`cor_test()` is used for comparing two continuous variables, with Pearson,
Kendall, or Spearman methods.

If Spearman method is chosen and either variable has a tie, the approximate
distribution is use in the `coin::spreaman_test()` function. This is usually the
preferred method over the asymptotic approximation, which is the method
`stats:cor.test()` uses in cases of ties. If the Spearman method is chosen and
there are no ties, the exact method is used from `stats:cor.test`.


```{r, echo=TRUE, warning=FALSE}

# Making Testing Dataset
cor_testing_data <- exampleData_BAMA %>%
  dplyr::filter(antigen %in% c('A1.con.env03 140 CF', 'B.MN V3 gp70'),
                visitno == 1,
                group == 1) %>%
  tidyr::pivot_wider(id_cols = pubID,
                     names_from = antigen,
                     values_from = magnitude)

stats::cor(cor_testing_data$`A1.con.env03 140 CF`,
           cor_testing_data$`B.MN V3 gp70`,
           method = 'spearman')

cor_test(
  x = cor_testing_data$`A1.con.env03 140 CF`,
  y = cor_testing_data$`B.MN V3 gp70`,
  method = 'spearman'
)

```


## Computing the Area Under a Curve (Integration Estimate) for Messy Data

`trapz_sorted()` is used when you are not presorting your data so that it is
strictly increasing along the x-axis or when there are 'NA' values that are not
removed prior to estimating the area under the curve.

```{r, echo=TRUE, warning=FALSE}
set.seed(93)
n <- 10
# unsorted data
x <- sample(1:n, n, replace = FALSE)
y <- runif(n, 0, 42)
pracma::trapz(x, y)
trapz_sorted(x,y)

# NA in data
x <- c(1:6, NA, 8:10)
y[2] <- NA
pracma::trapz(x, y)
trapz_sorted(x,y, na.rm = TRUE)

```

\clearpage

# Fancy Output Functions

These are functions designed to produce professional output that can easily be
printed in reports.

## P Values

`pretty_pvalues()` is used on p-values by rounding them to a specified number of
digits and using < for low p-values as opposed to scientific notation (i.e., "p
< 0.0001" if rounding to 4 digits), and by allowing options for emphasizing
p-values and specific characters for missing values.


```{r, echo=TRUE}

pvalue_example = c(1, 0.06753, 0.004435, NA, 1e-16, 0.563533)
# For simple p value display
pretty_pvalues(pvalue_example, digits = 3, output_type = 'no_markup')

# For display in report
table_p_Values <- pretty_pvalues(pvalue_example, digits = 3, 
                                 background = "yellow")
kableExtra::kable(
  table_p_Values,
  escape = FALSE,
  linesep = "",
  col.names = c("P-values"),
  caption = 'Fancy P Values'
) %>%
  kableExtra::kable_styling(font_size = 8.5,
                            latex_options = "hold_position")

```

You can also specify if you want `p=` pasted on the front of the p values, using
the `include_p` parameter.

\clearpage

## Basic Combining of Variables

`stat_paste()` is used to combine two or three statistics together, allowing
for different rounding and bound character specifications. Common uses for this
function are for:

- Mean (sd)
- Median [min, max]
- Estimate (SE of Estimate)
- Estimate (95% CI Lower Bound, Upper Bound)
- Estimate/Statistic (p value)

```{r, echo=TRUE}
# Simple Examples
stat_paste(stat1 = 2.45, stat2 = 0.214, stat3 = 55.3, 
           digits = 2, bound_char = '[')
stat_paste(stat1 = 6.4864, stat2 = pretty_pvalues(0.0004, digits = 3), 
           digits = 3, bound_char = '(')


exampleData_BAMA %>% 
  filter(visitno == 1) %>%  
  group_by(antigen, group) %>% 
  summarise(`Magnitude Info (Median [Range])` = 
              stat_paste(stat1 = median(magnitude), 
                         stat2 = min(magnitude),
                         stat3 = max(magnitude),
                         digits = 2, bound_char = '['),
            `Magnitude Info (Mean (SD))` = 
              stat_paste(stat1 = mean(magnitude), 
                         stat2 = sd(magnitude),
                         digits = 2, bound_char = '('))



```

\clearpage


## Advanced Combining of Variables

`paste_tbl_grp()` pastes together information, often statistics, from two groups.
There are two predefined combinations: mean(sd) and median[min,max], but user
may also paste any single measure together.

Example of summary information to be pasted together (partial output):

```{r, echo=FALSE, warning=FALSE, size='tiny'}

summary_info <- exampleData_BAMA %>%
  filter(visitno == 1) %>%
  mutate(group = paste0("Group", group)) %>%
  group_by(antigen, group) %>%
  summarise_at("magnitude",
               list(
                 n = length,
                 mean = mean,
                 sd = sd,
                 median = median,
                 min = min,
                 max = max
               )) %>%
  tidyr::pivot_longer(cols = n:max) %>%
  tidyr::unite(var, group, name) %>%
  tidyr::pivot_wider(names_from = var, values_from = value) %>%
  mutate(Group1 = "Group 1", Group2 = "Group 2")

summary_info %>%
  select(antigen,
         Group1_max,
         Group1_mean,
         Group1_median,
         Group1_min,
         Group1_n)

```


```{r, echo=TRUE, warning=FALSE}

summary_table <-  summary_info %>%
  paste_tbl_grp(
    vars_to_paste = c('n', 'mean_sd', 'median_min_max'),
    first_name = 'Group1',
    second_name = 'Group2'
  )

kableExtra::kable(
  summary_table,
  booktabs = TRUE,
  linesep = "",
  caption = 'Summary Information Comparison'
) %>%
  kableExtra::kable_styling(font_size = 6.5, latex_options = "hold_position") %>%
  kableExtra::footnote('Summary Information for Group 1 vs. Group 2, by Antigen and for Visit 1')

```



\clearpage

## Escape

`escape()` protects control characters in a string for use in a latex table or
caption.

```{r, echo=TRUE, warning=FALSE}

value_example <- c("testvalue", "test_value", "ampersand&")
escape(value_example)
escape("String_Entry %")

```


## Collapsing Rows

`collapse_group_row()` is for replacing repeated rows in a data.frame into NA
for nice printing. This is an alternative to `kableExtra::collapse_rows()`,
which does not work well for tables that span multiple pages (i.e. `longtable =
TRUE`).

```{r, echo=TRUE, warning=FALSE}
set.seed(341235432)
sample_df <- data.frame(
  x = c(1, 1, 1, 2, 2, 2, 2),
  y = c('test1', 'test1', 'test2', 'test1', 'test2', 'test2', 'test1'),
  z = c(1, 2, 3, 4, 5, 6, 7),
  outputVal = runif(7)
)

collapse_group_row(sample_df, x, y, z) %>%
  kableExtra::kable(
    booktabs = TRUE,
    linesep = "",
    row.names = FALSE
  ) %>%
  kableExtra::row_spec(3, hline_after = TRUE) %>% 
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

```

\clearpage

# Pairwise Comparison Functions

`pairwise_test_bin()`, `pairwise_test_cont()`, `cor_test_pairs()` are
functions that perform pairwise group (or any categorical variable) comparisons.
The function will go through all pairwise comparisons and output descriptive
statistics and relevant p values, calling the appropriate testing functions from
the [Testing and Estimate Functions](#testing-and-estimate-functions) section.

## Pairwise Comparison of Multiple Groups for a Binary Variable 

Simple example using `pairwise_test_bin`:

```{r, echo=TRUE, warning=FALSE, size = 'tiny'}

set.seed(1)
x_example <- c(NA, sample(0:1, 50, replace = TRUE, prob = c(.75, .25)),
  sample(0:1, 50, replace = TRUE, prob = c(.25, .75)))
group_example <- c(rep(1, 25), NA, rep(2, 25), rep(3, 25), rep(4, 25))

pairwise_test_bin(x_example, group_example) %>% 
  rename(Separation = 'PerfectSeparation')


```


Group comparison example using `pairwise_test_bin`:

```{r, echo=TRUE, warning=FALSE, results='asis'}

## Group Comparison
group_testing <- exampleData_ICS %>%
  filter(Population == 'IFNg Or IL2' & Group != 4) %>%
  group_by(Stim, Visit) %>%
  group_modify( ~ as.data.frame(
    pairwise_test_bin(x = .$response, group = .$Group,
                      method = 'barnard', alternative = 'less',
                      num_needed_for_test = 3, digits = 1,
                      latex_output = TRUE, verbose = FALSE
                      )
  )) %>%
  ungroup() %>%
  # Getting fancy p values
  mutate(
    ResponseTest = pretty_pvalues(
      ResponseTest, output_type = 'latex',
      sig_alpha = .1, background = 'yellow'
    )
  )
```

\clearpage

```{r, echo=TRUE, warning=FALSE, results='asis'}

kableExtra::kable( group_testing, escape = FALSE, booktabs = TRUE,
  linesep = "", caption = 'Response Rate Comparisons Across Groups'
) %>%
  kableExtra::kable_styling(
    font_size = 6.5,
    latex_options = c("hold_position", "scale_down")
    ) %>%
  kableExtra::collapse_rows(c(1:2),
                            row_group_label_position = 'identity',
                            latex_hline = 'full')
```

\clearpage

Time point comparison example (paired) using `pairwise_test_bin()`:


```{r, echo=TRUE, warning=FALSE, results='asis'}

## Timepoint Comparison
timepoint_testing <- exampleData_ICS %>%
  filter(Population == 'IFNg Or IL2' & Group != 4) %>% 
  group_by(Stim, Group) %>%
  group_modify(~ as.data.frame(
    pairwise_test_bin(x = .$response, group = .$Visit, id = .$pubID,
                      method = 'mcnemar', num_needed_for_test = 3, digits = 1,
                      latex_output = TRUE, verbose = FALSE))) %>% 
  ungroup() %>% 
  # Getting fancy p values
  mutate(ResponseTest = pretty_pvalues(ResponseTest, output_type = 'latex', 
                                       sig_alpha = .1, background = 'yellow'))


kableExtra::kable(timepoint_testing, escape = FALSE, booktabs = TRUE, 
                  caption = 'Response Rate Comparisons Across Visits') %>% 
  kableExtra::kable_styling(font_size = 6.5, 
                            latex_options = c("hold_position","scale_down")) %>% 
  kableExtra::collapse_rows(c(1:2), row_group_label_position = 'identity', 
                            latex_hline = 'full')

```



\clearpage

## Pairwise Comparison of Multiple Groups for a Continuous Variable 


Simple example using `pairwise_test_cont()`:

```{r, echo=TRUE, warning=FALSE}

set.seed(1)
x_example <- c(NA, rnorm(50), rnorm(50, mean = 5))
group_example <- c(rep(1, 25), rep(2, 25), rep(3, 25), rep(4, 25), NA)

pairwise_test_cont(x_example, group_example, digits = 1) %>% 
  rename(Separation = 'PerfectSeparation')


```

\clearpage

Group comparison example using `pairwise_test_cont()`:

```{r, echo=TRUE, warning=FALSE, results='asis'}

## Group Comparison
group_testing <- exampleData_ICS %>%
  filter(Population == 'IFNg Or IL2' & Group != 4) %>% 
  group_by(Stim, Visit) %>%
  group_modify(~ as.data.frame(
    pairwise_test_cont(x = pmax(.$PercentCellNet, 0.00001), group = .$Group,
                      method = 'wilcox', paired = FALSE, 
                      alternative = 'greater', num_needed_for_test = 3, 
                      digits = 3, log10_stats = TRUE))) %>% 
   ungroup() %>% 
  # Getting fancy p values
  mutate(MagnitudeTest = pretty_pvalues(
    MagnitudeTest, output_type = 'latex', 
    sig_alpha = .1, background = 'yellow')
  ) %>% 
  rename("Median (Range)" = Median_Min_Max, 'Mean (SD) (log10)' = log_Mean_SD)

kableExtra::kable(group_testing, escape = FALSE, booktabs = TRUE, 
                  caption = 'Magnitude Comparisons Across Groups') %>% 
  kableExtra::kable_styling(font_size = 6.5, 
                            latex_options = c("hold_position","scale_down")) %>% 
  kableExtra::collapse_rows(c(1:2), 
                            row_group_label_position = 'identity', 
                            latex_hline = 'full')
```

\clearpage

Time point comparison example (paired) using `pairwise_test_cont()`:


```{r, echo=TRUE, warning=FALSE, results='asis'}

## Timepoint Comparison
timepoint_testing <- exampleData_ICS %>%
  filter(Population == 'IFNg Or IL2' & Group != 4) %>% 
  group_by(Stim, Group) %>%
  group_modify(~ as.data.frame(
    pairwise_test_cont(x =  pmax(.$PercentCellNet, 0.00001), group = .$Visit, 
                       id = .$pubID, method = 'wilcox', paired = TRUE, 
                       num_needed_for_test = 3, 
                       digits = 3, log10_stats = TRUE))) %>% 
   ungroup() %>% 
  # Getting fancy p values
  mutate(MagnitudeTest = pretty_pvalues(
    MagnitudeTest, output_type = 'latex', 
    sig_alpha = .1, background = 'yellow')
    ) %>% 
  rename("Median (Range)" = Median_Min_Max, 'Mean (SD) (log10)' = log_Mean_SD)


kableExtra::kable(timepoint_testing, 
                  escape = FALSE, 
                  booktabs = TRUE, 
                  caption = 'Magnitude Comparisons Across Visits') %>% 
  kableExtra::kable_styling(font_size = 6.5, 
                            latex_options = c("hold_position","scale_down")) %>% 
  kableExtra::collapse_rows(c(1:2), 
                            row_group_label_position = 'identity', 
                            latex_hline = 'full')

```




\clearpage

## Pairwise Comparison of Continuous Variables (Correlation)


Simple example using `cor_test_pairs`:

```{r, echo=TRUE, warning=FALSE}

set.seed(1)
x_example <- c(1, 1, rnorm(48), rnorm(49, mean = 5), NA)
pair_example <- c(rep('antigen A', 25), rep('antigen B', 25), 
                   rep('antigen C', 25), rep('antigen D', 25))
id_example <- c(rep(1:25, 4))

cor_test_pairs(x_example, pair_example, id_example, digits = 4)


```

\clearpage

Correlation comparisons using `cor_test_pairs`:

```{r, echo=TRUE, warning=FALSE, results='asis'}
## Antigen Correlations
antigen_testing <- exampleData_BAMA %>%
  filter(antigen %in% c('B.63521 D11gp120/293F', 
                        'B.con.env03 140 CF',
                        'B.MN V3 gp70') & 
           group != 4) %>% 
  group_by(group, visitno) %>%
  group_modify(~ as.data.frame(
    cor_test_pairs(x = .x$magnitude,
                      pair = .x$antigen,
                      id = .x$pubID,
                      method = 'spearman',
                      n_distinct_value = 3))) %>% 
  ungroup() %>% 
  # Getting fancy p values
  mutate(CorrTest = pretty_pvalues(CorrTest, output_type = 'latex', 
                                   sig_alpha = .1, background = 'yellow')) %>% 
  rename("Corr P Value" = CorrTest)

kableExtra::kable(antigen_testing, escape = FALSE, 
                  booktabs = TRUE, caption = 'Correlations between Antigen') %>% 
  kableExtra::kable_styling(font_size = 6.5, 
                            latex_options = c("hold_position","scale_down")) %>% 
  kableExtra::collapse_rows(c(1:2), 
                            row_group_label_position = 'identity', 
                            latex_hline = 'full')
```


\clearpage


# Survival and Magnitude Breadth

`create_step_curve()` creates survival probabilities from time and censoring
information and generates a risk table that includes the survival probabilities
and number at risk in addition to the data provided. This output can be used to
plot step line outcomes such as time-to-event (Kaplan-Meier curves), magnitude
breadth (MB) curves, and cumulative plots.


```{r, echo=TRUE, warning=FALSE, fig.cap='Potency-breath curves'}

#Potency-breadth curves
plot_data <-
 CAVD812_mAB %>%
  filter(virus != 'SVA-MLV') %>% 
  tidyr::pivot_longer(cols = c(ic50, ic80)) %>% 
  dplyr::group_by(name, product) %>%
  dplyr::group_modify(~ create_step_curve(x = pmin(.x$value, 100), 
                                          event = as.numeric(.x$value < 50),
                                          flip_surv = TRUE, flip_top_x = 100)) 
  
  ggplot2::ggplot(data = plot_data,
                  ggplot2::aes(x = time, y = 1 - surv, color = product)) +
  ggplot2::geom_step(direction = 'hv', lwd = .35) +
  ggplot2::scale_x_log10() +
  ggplot2::scale_y_continuous('Viral Coverage (%)') +
  ggplot2::facet_grid(. ~ name) +
  ggplot2::theme_bw()
```

\clearpage

`mb_results()` creates step curve info for magnitude breadth (MB) plots, with
options to include response status and have logged transformation for AUC-MB
calculation.

```{r, echo=TRUE, warning=FALSE, fig.cap='BAMA Magnitude Breadth curves'}
data_here <- exampleData_BAMA %>% filter(visitno == 2)

group_results <- data_here %>% dplyr::group_by(group) %>%
  dplyr::group_modify(~ mb_results(magnitude = .x$magnitude , 
                                   response = .x$response))
ind_results <-  data_here %>% dplyr::group_by(group, pubID) %>%
  dplyr::group_modify(~ mb_results(magnitude = .x$magnitude , 
                                   response = .x$response))
ggplot2::ggplot(
  data = group_results,
  ggplot2::aes(x = magnitude, y = breadth, color = factor(group))) +
  ggplot2::geom_step(data = ind_results, ggplot2::aes(group = pubID),
                     linetype = "dashed", direction = 'hv', lwd = .35, alpha = .7) +
  ggplot2::geom_step(direction = 'hv', lwd = .65) +
  ggplot2::scale_x_log10('Response Magnitude',
    breaks = c(100,1000,10000, 22000),
    labels = c(expression("" <= 100),1000,10000, expression("" >= 22000))) +
  ggplot2::ylab('Magnitude Breadth (%)') +
  ggplot2::scale_color_discrete('Group') +
  ggplot2::coord_cartesian(xlim = c(95, 23000)) +
  ggplot2::theme_bw()
```

\clearpage

```{r, echo=TRUE, warning=FALSE, fig.cap='BAMA Magnitude Breadth curves'}

# AUC-MB plot
AUC_MB <- dplyr::distinct(ind_results, group, pubID, aucMB)

ggplot2::ggplot(AUC_MB, ggplot2::aes(x = factor(group), y = aucMB,
                                     color = factor(group))) +
  ggplot2::geom_boxplot(outlier.color = NA, show.legend = FALSE) +
  ggplot2::geom_point(position = ggplot2::position_jitter(width = 0.25,
           height = 0, seed = 1), size = 1.5, show.legend = FALSE) +
  ggplot2::scale_y_log10('AUC-MB') +
  ggplot2::xlab('Group') +
  ggplot2::theme_bw()
```


\clearpage


# Utility Functions


`round_away_0()` is a function to properly perform mathematical rounding (i.e.
rounding away from 0 when tied), as opposed to the `round()` function, which
rounds to the nearest even number when tied. Also `round_away_0()` allows for
trailing zeros (i.e. 0.100 if rounding to 3 digits).

```{r}
vals_to_round = c(NA,-3.5:3.5)
vals_to_round
round(vals_to_round)
round_away_0(vals_to_round)
round_away_0(vals_to_round, digits = 2, trailing_zeros = TRUE)
```

\clearpage

`get_session_info()` produces reproducible tables, which are great to add to the
end of reports. The first table gives software session information and the
second table gives software package version information. `get_full_name()` is a
function used by `get_session_info()` to get the user's name, based on user's ID.

```{r}
my_session_info <- get_session_info()
kableExtra::kable(
  my_session_info$platform_table,
  booktabs = TRUE,
  linesep = '',
  caption = "Reproducibility Software Session Information"
) %>%
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

```

```{r}

kableExtra::kable(
  my_session_info$packages_table,
  booktabs = TRUE,
  linesep = '',
  caption = "Reproducibility Software Package Version Information"
) %>%
  kableExtra::kable_styling(font_size = 8, latex_options = "hold_position")

```
