---
title: "Plotting utilities in VISCfunctions"
author: "Kellie MacPhee"
date: "`r Sys.Date()`"
output: 
  html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE
)
```

In this vignette, we will overview the following elements of VISCfunctions,
which help with plotting:

- create_step_curve (for survival and magnitude breadth analysis)

## Getting Started

Install and load the VISCfunctions package, along with other packages needed
for this vignette:

```{r, eval=FALSE}
remotes::install_github("FredHutch/VISCfunctions")
```

```{r, results='hide', message=FALSE}
library(VISCfunctions)
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Plotting

`create_step_curve()` creates survival probabilities from time and censoring
information and generates a risk table that includes the survival probabilities
and number at risk in addition to the data provided. This output can be used to
plot step line outcomes such as time-to-event (Kaplan-Meier curves), magnitude
breadth (MB) curves, and cumulative plots.

```{r}
plot_data <- CAVD812_mAB %>%
  filter(virus != 'SVA-MLV') %>% 
  pivot_longer(cols = c(ic50, ic80)) %>% 
  group_by(name, product) %>%
  group_modify(~ create_step_curve(x = pmin(.x$value, 100), 
                                   event = as.numeric(.x$value < 50),
                                   flip_surv = TRUE, flip_top_x = 100)) 

plot_data %>%
  ggplot(aes(x = time, y = 1 - surv, color = product)) +
  geom_step(direction = 'hv', lwd = .35) +
  scale_x_log10() +
  scale_y_continuous('Viral Coverage (%)') +
  facet_grid(. ~ name) +
  theme_bw()
```
```{r}

scale_y_continuous_with_limits <- function(lower_limit = NA,
                                           upper_limit = NA,
                                           target_breaks = 3,
                                           breaks = NULL,
                                           oob = scales::squish,
                                           ...) {
  
  if (is.null(breaks)) {
    if (is.na(lower_limit) & is.na(upper_limit)) {
      my_breaks <- waiver()
    }
    if (!is.na(lower_limit) & !is.na(upper_limit)) {
      width <- upper_limit - lower_limit
      my_breaks <- seq(lower_limit, upper_limit, width/(target_breaks - 1))
    }
    if (!is.na(lower_limit) & is.na(upper_limit)) {
      break_spacing <- 10^round(log10(lower_limit))
      my_breaks <- lower_limit + c(0, break_spacing * 1:(target_breaks - 1))
    }
    if (is.na(lower_limit) & !is.na(!upper_limit)) {
      break_spacing <- round(upper_limit / target_breaks)
      my_breaks <- upper_limit - c(break_spacing * (target_breaks - 1):1, 0)
    }
  } else {
    my_breaks <- breaks
  }
  
  if (!is.na(lower_limit)) stopifnot(lower_limit %in% my_breaks)
  if (!is.na(upper_limit)) stopifnot(upper_limit %in% my_breaks)
  
  my_labeller <- function(breaks) {
    labels <- as.character(breaks)
    n_labels <- length(labels)
    if (!is.na(lower_limit)) { labels[1] <- paste("≤", labels[1]) }
    if (!is.na(upper_limit)) { labels[n_labels] <- paste("≥", labels[n_labels]) }
    return(labels)
  }
  
  ggplot2::scale_y_continuous(
    limits = c(lower_limit, upper_limit),
    breaks = my_breaks,
    labels = my_labeller,
    oob = oob,
    ...
  )
  
}


mtcars %>%
  ggplot(aes(wt, mpg)) +
  geom_point() +
  scale_y_continuous_with_limits(lower_limit = 12)

mtcars %>%
  ggplot(aes(wt, mpg)) +
  geom_point() +
  scale_y_continuous_with_limits(upper_limit = 30, target_breaks = 4)

mtcars %>%
  ggplot(aes(wt, mpg)) +
  geom_point() +
  scale_y_continuous_with_limits(lower_limit = 12, upper_limit = 30, target_breaks = 4)

mtcars %>%
  ggplot(aes(wt, mpg)) +
  geom_point() +
  scale_y_continuous_with_limits()
```
